### 2025-03-23 article/article.qmd
cleanup

---
title: CLEM-ICD - Clinical Language Encoding with ModernBERT
subtitle: Using ModernBERT to Improve Automated ICD-10 Classification
author:
  - name: Tyler Cross
    email: tyler.cross@berkeley.com
    affiliations: 
        - id: some-u
          name: University of California, Berkeley
          department: School of Information
          address: 102 South Hall
          city: Berkeley
          state: CA
          postal-code: 94720
    attributes:
        corresponding: true
    note: |
      Final project submission for the UC Berkeley Master of Information and Data Science program, DATASCI 266: Natural Language Processing with Deep Learning
abstract: |
  Medical coding—the assignment of standardized ICD-10 codes to clinical documentation—remains a labor-intensive process requiring expert manual review of extensive narratives against 150,000+ potential classifications. We present CLEM-ICD, an extension of the PLM-ICD framework that replaces RoBERTa with ModernBERT to leverage its 8192-token context window for improved automated medical coding. Our architectural enhancement preserves the multi-label classification paradigm while significantly expanding the model's capacity to process lengthy clinical text, capturing distant dependencies crucial for accurate code assignment. Experiments on the MIMIC-IV dataset demonstrate that CLEM-ICD achieves superior performance compared to previous approaches, as measured by precision, recall, and F1-score. This work addresses a critical bottleneck in healthcare administration through advanced NLP techniques, offering a scalable solution for reducing the cognitive burden of medical coding while maintaining diagnostic accuracy.
keywords: 
  - clinical natural language processing
  - medical coding automation
  - transformer models
  - multi-label classification
  - long-context language models
  - ModernBERT
  - MIMIC-IV
  - ICD-10
  - healthcare informatics
  - biomedical text classification
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      # model: 3p # Don't set a model with preprint
      cite-style: authoryear
---

Medical coding is a critical yet complex process in healthcare administration. The translation of clinical documentation into standardized ICD-10 codes is essential for healthcare reimbursement, clinical research, and public health statistics. Currently, healthcare providers rely on dedicated medical coders who manually review extensive clinical notes to assign appropriate codes from over 150,000 possible diagnostic and procedural codes, creating significant administrative overhead.

For every $1 of revenue collected by a hospital, approximately $0.25 is spent on administrative tasks necessary to collect it. Non-doctor workers outnumber doctors in healthcare roughly 16 to 1, with medical coding representing a significant portion of this burden.

This project aims to improve automated medical coding by leveraging ModernBERT's enhanced context window capabilities, potentially reducing the administrative burden while maintaining or improving coding accuracy.

# Background

@edin2024explainable recently demonstrated success in medical coding using a RoBERTa-based approach. Their work showed promising results but left room for improvement in both performance and contextual understanding. Other notable contributions in this field include [additional related work to be filled in].

ModernBERT [@warner2024modernbert] represents an advancement over previous BERT-like models, offering an enhanced context window of 8192 tokens compared to the 512 tokens in traditional models. This increased context capacity is particularly valuable for medical documents, which tend to be lengthy and contain important information distributed throughout the text.

## Using CSL

# Methods

This project utilizes the MIMIC-IV dataset, a large, freely available database comprising de-identified health data associated with hospital stays. The dataset includes detailed clinical notes, diagnostic codes, procedural information, and other health-related data from real hospital encounters.

## System Architecture

Our approach replaces the RoBERTa model used in previous work [@edin2024explainable] with ModernBERT [@warner2024modernbert], leveraging its enhanced context window to better process lengthy clinical notes. We hypothesize that this will lead to improved code prediction by enabling the model to capture more comprehensive context from medical documents.

## Implementation Details

[To be completed with specific implementation details]

## Evaluation Approach

We evaluate our system using standard performance metrics: precision, recall, F1-score, and accuracy compared to gold-standard human coding, following the evaluation protocol established in previous work.

# Results and Discussion

[This section will be completed after obtaining experimental results]

# References {-}